A Plasma Torus Around a Young Low-Mass Star

### 2025-01-21 paper/ms.tex
update

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars that host which can trap outflowing stellar winds into
warped plasma tori \cite{Townsend2005,Townsend2008}.  These tori tend
to have dense antipodal accumulations of plasma sculpted by
tilted-dipole magnetic fields, and the transits of
these clumps are thought to produce the observed broadband photometric variability
through a combination of bound-free scattering \cite{Townsend2005} and
Thomson scattering \cite{Berry2022}.  For $\sigma$~Ori~E and almost
all of its analogs, the result is light curves that appear ``simple'',
resembling those of eclipsing binaries.  The two known
exceptions, HD~37776 and HD~64740, show complex light curves
resembling CPVs \cite{Mikulasek2020,Bouma2024} and have
spectropolarimetric magnetic field maps indicating strong
contributions from higher-order magnetic moments
\cite{Kochukhov2011,Shultz2018}.  There are two implications: firstly,
the complexity of CPVs may be a direct consequence of magnetic fields
with highly multipolar contributions.  Secondly, CPVs could be a
source of astrophysical false positives in photometric searches for
eclipsing binaries and transiting exoplanets around young
pre-main-sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The role of the star's magnetic field could be better understood
through new observations, and new theory.  From the theoretical
perspective, there is an urgent need for rigid-field
(magneto)-hydrodynamic modeling to go beyond previous work
\cite{Townsend2005,Townsend2008} and to explore the effects of
non-dipolar field contributions.
Fully time-dynamic MHD \cite{Daley-Yates2024} will offer the
capability of understanding the connection between the plasma and the
dust.  %TODO: need this sentence to be better
Observationally, optical spectropolarimetry has the potential to
assess both the field strength and topology.  A more direct probe
however might be to connected the recent work \cite{Kaur2024} showing
that CPVs are variable radio emitters, exhibiting emission components
that can be both persistent, as well as short-lived and highly
polarized.  This opens the prospects for detecting for radio waves
emitted through the electron cyclotron maser instability, which can
provide a direct measurement of the field strength at the site of the
emitting region.

@@

\subsection{Stellar Parameters}
{\it Age}
While most CPVs can be spatially and kinematically associated with known young
moving groups using tools such as BANYAN\,$\Sigma$ \cite[v1.2;][]{Gagne2018},
TIC~141146667 cannot \cite{Bouma2024}.
%.\footnote{\url{https://github.com/jgagneastro/banyan_sigma},
%git commit \texttt{394b486}}

Isochrone age: $\sim$TucHor or younger.  Maybe 30-35myr?
NB color puts it at ~M4.5 (main seq).
BP-RP=3.3..

Lithium?
Frankly, it's not obviously there!  (but maybe this is bc of the blending, and
a pEW would show smth small)
% We have K=10.47, d=57pc - > DM ~= 3.78 -> M_K ~= 6.67
Figure7 of Wood23 suggests from the Feiden Li depletion curves that no lithium would imply $t$>37Myr.
However Figure5 frmo the same work suggests maybe try binning up the spectrum, and then comparing against Bochanski+2007 for a lithium-free template.

>>>>

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars that host which can trap outflowing stellar winds into
warped plasma tori \cite{Townsend2005,Townsend2008}.  These tori tend
to have dense antipodal accumulations of plasma sculpted by
tilted-dipole magnetic fields, and the transits of
these clumps are thought to produce the observed broadband photometric variability
through a combination of bound-free scattering \cite{Townsend2005} and
Thomson scattering \cite{Berry2022}.  For $\sigma$~Ori~E and almost
all of its analogs, the result is light curves that appear ``simple'',
resembling those of eclipsing binaries \cite{Townsend2008}.  The two known
exceptions, HD~37776 and HD~64740, show complex light curves
resembling CPVs \cite{Mikulasek2020,Bouma2024} and have
spectropolarimetric magnetic field maps indicating strong
contributions from higher-order magnetic moments
\cite{Kochukhov2011,Shultz2018}.  There are two implications: firstly,
the complexity of CPVs may be a direct consequence of magnetic fields
with highly multipolar contributions.  Secondly, CPVs could be a
source of astrophysical false positives in photometric searches for
eclipsing binaries and transiting exoplanets around young
pre-main-sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The role of the star's magnetic field could be better understood
through new observations, and new theory.  From the theoretical
perspective, there is an urgent need for rigid-field
(magneto)-hydrodynamic modeling to go beyond previous work
\cite{Townsend2005,Townsend2008} and to explore the effects of
non-dipolar field contributions.
In particular, dynamo simulations of fully-convective M dwarfs have
suggested the possibility that global-scale mean fields can be
confined to a single hemisphere \cite{Brown2020}; such fields would
significantly alter plausible accumulations of circumstellar material
from those that have been previously explored \cite{Townsend2008}.
Observationally, optical spectropolarimetry has the potential to
assess both the field strength and topology.  A more direct probe
however might be to connected the recent work \cite{Kaur2024} showing
that CPVs are variable radio emitters, exhibiting emission components
that can be both persistent, as well as short-lived and highly
polarized.  This opens the prospects for detecting for radio waves
emitted through the electron cyclotron maser instability, which can
provide a direct measurement of the field strength at the site of the
emitting region.

@@

\subsection{Stellar Parameters}
{\it Age: No Obvious Association Membership}---Previous work
  \cite{Bouma2024} has shown that an overwhelming majority of CPVs can
  be associated with known young moving groups based on their
  positions and kinematics using probabilistic tools such as
  BANYAN\,$\Sigma$ \cite[v1.2;][]{Gagne2018}.  TIC~141146667 cannot;
  based on the current census of these known groups, we calculated its
  probability of being part of any of these groups using its known
  three-dimensional galactic position velocity, and it appears to be a
  field star at $>$99.9\% confidence.  Searching the local vicinity of
  the star for neighbors with similar projected on-sky velocity (CITE)
  similarly yields no strong candidates for co-moving stars with
  tangential velocity differences below 5\,\kms\ that share its
  isochronal youth.

{\it Age: Isochrones}--- $\sim$TucHor or younger.  Maybe 30-35myr?  NB
  color puts it at ~M4.5 (main seq).  BP-RP=3.3..
    % uhhh careful w what you mean by "isochrone"

{\it Age: Lithium}--- Frankly, it's not obviously there!  (but maybe
this is bc of the blending, and a pEW would show smth small)
% We have K=10.47, d=57pc - > DM ~= 3.78 -> M_K ~= 6.67
Figure7 of Wood23 suggests from the Feiden Li depletion curves that no lithium would imply $t$>37Myr.
However Figure5 frmo the same work suggests maybe try binning up the spectrum, and then comparing against Bochanski+2007 for a lithium-free template.
